---
title: "Create main data frame"
date: "`r Sys.Date()`"
output: html_document
params:
  save_to_disk: FALSE
---



# Setup

```{r message = FALSE}
library(tidyverse)
library(sjmisc)
library(here)
```


Where am I?

```{r}
here::here()
```


Load raw data:

```{r}
rawdata <- readxl::read_xlsx(paste0(here::here(), "/raw-data/Rohdaten.xlsx"))
```


Add id:

```{r}
rawdata <- rawdata %>% 
  add_id(var = "id")
```


Check names:

```{r}
names(rawdata)[c(30, 31)]
```


Check level-1 (l1) data (reaction times):
```{r}
head(rawdata$NW03_01)
```


## Extract l1 data



test it:
```{r}
rawdata$NW03_01[1] %>% str_split(pattern=",") %>% `[[`(1) %>% parse_number() %>% str()
```

### block 1: "flip as often as you can"

Extract reaction time (rt) data, block 1 (flip as often as you can):

```{r}
flip_rt <- rawdata$NW04_01 %>% 
  map(~str_split(., pattern = ",")) %>% 
  map(~`[[`(., 1)) %>%  # extract string element from list, not list, as parse_number needs string element
  map(~ parse_number(.))
```


Get descriptives:

```{r}

flip_stats <- list()

flip_stats$md_flip <- flip_rt %>% flatten_dbl() %>% median(na.rm = T)
flip_stats$mean_flip <- flip_rt %>% flatten_dbl() %>% mean(na.rm = T)
flip_stats$sd_flip <- flip_rt %>% flatten_dbl() %>% sd(na.rm = T)
flip_stats$iqr_flip <- flip_rt %>% flatten_dbl() %>% median(na.rm = T)


flip_stats


```


Maybe better as as tibble:

```{r}

flip_rt_sum <- flip_stats %>% 
  flatten_dbl() %>% 
  enframe(name = "trial", value = "rt") %>% 
  drop_na() %>% 
  summarise(type = "flip",
            md_rt = median(rt),
            mean_rt = mean(rt),
            sd_rt = sd(rt),
            iqr_rt = IQR(rt)
            )


flip_rt_sum
```



Visualize it:

```{r}
flip_rt %>% 
  flatten_dbl() %>% 
  enframe() %>% 
  ggplot(aes(x = value)) +
    geom_histogram() +
  geom_vline(xintercept = flip_stats$md_flip, 
             linetype = "dashed") +
  labs(caption = "The dashed line shows the median",
       x = "Reaction time [ms]",
              title = "Reaction time distribution for block flip") +
  annotate("label", y = 0, x = flip_stats$md_flip, 
           label = round(flip_stats$md_flip))
```


### block 2: "hold as long as you can"



```{r}
hold_rt <- rawdata$NW03_01 %>% 
  map(~str_split(., pattern = ",")) %>% 
  map(~`[[`(., 1)) %>% 
  map(~ parse_number(.))
```



Get descriptives:

```{r}

hold_rt_sum <- hold_rt %>% 
  flatten_dbl() %>% 
  enframe(name = "trial", value = "rt") %>% 
  drop_na() %>% 
  summarise(type = "hold",
            md_rt = median(rt),
            mean_rt = mean(rt),
            sd_rt = sd(rt),
            iqr_rt = IQR(rt)
            )


hold_rt_sum
```




Visualize it:

```{r}


hold_rt %>% 
  flatten_dbl() %>% 
  enframe() %>% 
  ggplot(aes(x = value)) +
    geom_histogram() +
  geom_vline(xintercept = hold_rt_sum$md_rt[1], 
             linetype = "dashed") +
  labs(caption = "The dashed line shows the median",
       x = "Reaction time [ms]",
       title = "Reaction time distribution for block FLIP") +
  annotate("label", y = 0, x = hold_rt_sum$md_rt[1], 
           label = round(hold_rt_sum$md_rt[1]))
```



# Shape to very long data frame



## Build rt data frame

Build reaction time (rt) data frame, where each flip (of each person) is a row


Name list elements:

```{r}
name_list_elements <- paste0("id_",1:143)

names(flip_rt) <- name_list_elements
names(hold_rt) <- name_list_elements
```


### Flip data frame
```{r}
rt_flip_long <- enframe(flip_rt) %>% 
  unnest() %>% 
  add_column(type = "flip")

rt_flip_long
```


### Hold data frame

```{r}
rt_hold_long <- enframe(hold_rt) %>% 
  unnest() %>% 
  add_column(type = "hold")

rt_hold_long
```


### Merge hold and flip data frames


```{r}
rt_long <- rt_flip_long %>% 
  bind_rows(rt_hold_long)
```


Get ID variable back:

```{r}
rt_long2 <- rt_long %>% 
  separate(name, into = c("dummy", "id"), sep = "_") %>% 
  remove_cols(dummy) %>% 
  mutate(id = as.numeric(id)) %>% 
  rename(rt = value)

rt_long2
```


# Build master data frame on level 1 data

Each row is a flip, so a very long data frame.

This type of data frame is built because (I assume) modeling functions can only digest data where a row is a level 1 data (ie., a flip)


```{r}
rawdata_long <- rawdata %>% 
  full_join(rt_long2)
```

Is the length of the resulting data frame equal to the length of the RT data frame?

```{r}
nrow(rawdata_long) == nrow(rt_long2)
```



# Compute mindfulness score (as MEAN)


```{r}
rawdata_long <- rawdata_long %>% 
  row_means(MA01_01:MA01_17, n = .9,
            var = "mindlessness_score")

hist(rawdata_long$mindlessness_score)
```

How many NAs?

```{r}
sum(is.na(rawdata_long$mindlessness_score))
```

Recode mindfulness score (originally, it is code as mindlessness):

```{r}
rawdata_long2 <- rawdata_long %>% 
  mutate(mindfulness_score = 7 - mindlessness_score)
```


# Rename variables


## Mindfulness training experience


Does the person trains mindfulness (regularly)?

```{r}
rawdata_long3 <- rawdata_long2 %>% 
  mutate(mindfulness_training = ifelse(PE06 == 1, 1, 0))


```


Training YES: 1
Training NO: 0




# Collapse data on person-level


Remove level 1 data:

```{r}
d_person <- select(rawdata_long3, -c(rt, type)) %>% 
  distinct()

d_person
```


Summarise level 1 data on person level and on type (block) level, that is, on level 2.


block HOLD

```{r}
rt_summary_hold <- rawdata_long3 %>% 
  group_by(id, type) %>% 
  filter(type == "hold") %>% 
  summarise(rt_md_hold = median(rt, na.rm = TRUE),
            rt_mean_hold = mean(rt, na.rm = TRUE),
            rt_iqr_hold = IQR(rt, na.rm = TRUE),
            rt_sd_hold = IQR(rt, na.rm = TRUE))  %>% 
  ungroup() %>% 
  select(-type)

rt_summary_hold
```



block FLIP

```{r}
rt_summary_flip <- rawdata_long3 %>% 
  group_by(id, type) %>% 
  filter(type == "flip") %>% 
  summarise(rt_md_flip = median(rt, na.rm = TRUE),
            rt_mean_flip = mean(rt, na.rm = TRUE),
            rt_iqr_flip = IQR(rt, na.rm = TRUE),
            rt_sd_flip = IQR(rt, na.rm = TRUE)) %>% 
  ungroup() %>% 
  select(-type)

rt_summary_flip
```



Join the summary data frames:


```{r}
rt_summary <- rt_summary_hold %>% 
  full_join(rt_summary_flip) %>% 
  ungroup()

rt_summary
```







Add rt summaries to person/type level (level 2) data frame:


```{r}
d_person_rt_sum <- d_person %>% 
  full_join(rt_summary) %>% 
  select(id, contains("rt"), mindfulness_score, mindfulness_training, everything() )

d_person_rt_sum
```






# Save data to disk



```{r echo = FALSE, eval = params$save_to_disk}
save(rawdata, file = paste0(here::here(), "/raw-data/raw_data_person_level.Rdata"))
save(d_person_rt_sum, file = paste0(here::here(), "/raw-data/raw_data_person_level_rt_summary.Rdata"))
save(rawdata_long3, file = paste0(here::here(), "/objects/rawdata_long.Rdata"))

cat("Data files have been saved to disk.")
```


```{r echo = FALSE, eval = !params$save_to_disk}

cat("Data files have NOT been saved to disk.")
```





# SessionInfo



```{r}
sessionInfo()
```

