---
title: "Summary stats on RAW data"
date: "`r Sys.Date()`"
output: html_document
number_sections: TRUE
---



# Setup


## Libs

```{r message = FALSE}
library(tidyverse)
library(sjmisc)
library(here)
library(dabestr)
```


## Where am I?

```{r}
here::here()
```


## Load data

```{r}
load(file = paste0(here::here(), "/objects/rawdata_long.Rdata"))

d <- rawdata_long3

d <- d %>% 
  select(rt, type, id, mindfulness_score, mindfulness_training, everything())
d
```


## Data structure


There are 3 hierarchical levels in the data:

- level 1: trials, single flip response
- level 2: blocks, encompassing several flips; there are 2 blocks
- level 3: person, encompassing 2 blocks, there are 143 (raw) persons








# Summary RT data


## Grand RT distribution

```{r}
d %>% 
  select(rt) %>% 
  ggplot(aes(x  = rt)) +
  geom_histogram(bins = 30) +
  geom_vline(xintercept = median(d$rt, na.rm = T), 
             linetype = "dashed") +
  labs(caption = "The dashed vertical line shows the median") +
  annotate("label", x = median(d$rt, na.rm = T),
           y = 0, label = median(d$rt, na.rm = T))
```

## RT summaries per block (`type`)


```{r}

d_sum <- d %>%
  select(rt, type) %>% 
  drop_na() %>% 
  group_by(type) %>% 
  summarise(rt_md = median(rt, na.rm = T),
            rt_sd = sd(rt),
            rt_iqr = IQR(rt),
            rt_mean = mean(rt))
 d_sum 
```


##  RT distribution per block


```{r}

 d %>% 
  select(rt, type) %>% 
  ggplot(aes(x  = rt)) +
  facet_wrap(~ type, scales = "free") +
  geom_histogram(bins = 30) +
  geom_vline(xintercept = median(d$rt, na.rm = T), 
             linetype = "dashed") +
  labs(caption = "The dashed vertical line shows the median") +
  geom_label(data = d_sum,
             aes(x = rt_md, label = rt_md),
             y = 0)
```


## RT per person per block

Let's aggregate on the person level, and compare then the blocks


```{r}
 d_sum2 <- d %>% 
  select(rt, type, id) %>%
  drop_na() %>% 
  group_by(id, type) %>% 
  summarise(rt_id_md = median(rt),
            rt_id_mean = mean(rt))

d_sum2
```


```{r}
d_sum2 %>% 
  ggplot(aes(x = type, y = rt_id_md)) +
  geom_boxplot() +
  geom_jitter(width = .1, alpha = .3) +
  scale_y_log10()
```


## RT (collapsed on l2) by mindfulness score


```{r}
d %>% 
  select(rt, type, id, mindfulness_score) %>% 
  group_by(type, id, mindfulness_score) %>% 
  summarise(rt = median(rt)) %>% 
  ggplot(aes(x = mindfulness_score, y = rt, color = type)) +
  geom_point() +
  scale_y_log10() +
  geom_smooth()
```

No pattern discernible, I'd say.

MAYBE there's a pattern in the middle of the mindfulness score range, but surely not at the edges.


## RT collapsed b mindfulness training


```{r}
d %>% 
  select(rt, type, id, mindfulness_training) %>% 
  group_by(type, id, mindfulness_training) %>% 
  summarise(rt = median(rt)) %>% 
  ggplot(aes(x = factor(mindfulness_training), y = rt, color = type)) +
  geom_boxplot() +
  geom_jitter(alpha = .3) +
  scale_y_log10() +
  facet_grid(~type)
```



It appears that mindfulness practitioners can exert a stronger attentional control: When asked to flip, they flip quicker. When asked to hold, they hold longer (in comparion to non-mindfulness-practitioners).



## RT Distribution by person by block


```{r}
d %>% 
  select(rt, type, id) %>% 
  group_by(type) %>% 
  ggplot(aes(x = factor(id), y = log10(rt))) +
  geom_boxplot() +
  facet_wrap(~ type, nrow = 2) +
  labs(x = "Persons")
```



# Summary mindfulness data


## Collapse main data frame 

So that one individual (`id`) matches to one row.

Only `mindfulness_training`, `mindfulness_score` and `id` are used.

```{r}
d_person <- d %>% 
  group_by(mindfulness_training, id) %>% 
  summarise(mindfulness_score = median(mindfulness_score)) %>% 
  ungroup() %>% 
  mutate(mindfulness_training_f = factor(mindfulness_training))

d_person

unique(d_person$mindfulness_training_f)
```


## Mindfulness practitioners show a higher mindfulness score

... compared to non-practitioners





```{r}
d_person %>% 
  ggplot(aes(x = factor(mindfulness_training), 
             y = mindfulness_score)) +
  geom_jitter(width = .1, alpha = .3) +
  geom_boxplot(alpha = .7) 
```


There appears to be an effect in that mindfulness practitioniers describe themselves as more mindful (compared to the self-description on persons not practicing mindfulness).




```{r eval = FALSE}

# CURRENTLY NOT WORKING, BUG?

unpaired_mean_diff <- dabest(iris, Species, Petal.Width,
                             idx = c("setosa", "versicolor"),
                             paired = FALSE)

dabest <- dabest(
  .data = d_person, 
  x = mindfulness_training_f, 
  y = mindfulness_score,
  idx = c("0", "1"),
  paired = FALSE,
  func = median
)
```












# SessionInfo



```{r}
sessionInfo()
```